%!TEX root = fastZKP.tex
\section{Preliminaries}
\label{sec::prelim}

\subsection{Notation}

In this paper, we use $\lambda$ to denote the security parameter, and $\neg(\lambda)$ to denote the negligible function in $\lambda$. ``PPT" stands for probabilistic polynomial time. We use $f(),h()$ for polynomials, $x,y,z$ for vectors of variables and $g,u,v$ for vectors of values. $x_i$ denotes the $i$-th variable in $x$. We use bold letters such as $\textbf{A}$ to represent arrays. For a multivariate polynomial $f$, its "variable-degree" is the maximum degree of $f$ in any of its variables.

\paragraph{Bilinear pairings.} Let $\mathbb{G}, \mathbb{G}_T$ be two groups of prime order $p$ and let $g\in\mathbb{G}$ be a generator. $e: \mathbb{G}\times\mathbb{G}\rightarrow\mathbb{G}_T$ denotes a bilinear map and we use $\mathsf{bp}=(p,\mathbb{G},\mathbb{G}_T,e,g)\leftarrow\mathsf{BilGen}(1^\lambda)$ for the generation of parameters for the bilinear map. Our scheme relies on the $q$-Strong Bilinear Diffie-Hellman ($q$-SBDH) assumption and an extended version of the Power Knowledge of Exponent (PKE) assumption. 


\begin{assumption}[$q$-Strong Bilinear Diffie-Hellman]
	\label{asp::qSDH}
	For any PPT adversary $\mathcal{A}$, the following holds:
	
	\[\Pr\left[ \begin{aligned}
	&\mathsf{bp} \leftarrow \mathsf{BilGen}(1^\lambda) & \\
	&s \overset{R}{\leftarrow} \mathbb{Z}^{*}_{p} & : (x, e(g, g)^{\frac{1}{s+x}}) \leftarrow \mathcal{A}(1^\lambda, \sigma)\\
	&\sigma = (\mathsf{bp}, g^s, ..., g^{s^q})
	\end{aligned} \right] \le \negl{(\lambda)}\]
\end{assumption}

The second assumption is a generalization of the $q$-PKE assumption~\cite{groth2010short} to multivariate polynomials, proposed in~\cite{zhang2017vsql,zkvpd}. Let $\mathcal {W}_ {\ell,d}$ be the set of all multisets of $\{1, . . . , \ell\}$ with the cardinality of each element being at most $d$. 

\begin{assumption}[$(d,\ell)$-Extended Power Knowledge of Exponent]
	\label{asp::dlEPKE}
	For any PPT adversary $\mathcal{A}$, there is a polynomial time algorithm $\mathcal{E}$ (takes the same randomness of $\mathcal{A}$ as input) such that for all benign auxiliary inputs $z \in \{0,1\}^{\poly({\lambda})}$ the following probability is negligible:
	{
		\[\Pr\left[ \begin{aligned}
		\mathsf{bp} \leftarrow \mathsf{BilGen}(1^\lambda) && \\
		s_1, ..., s_\ell, s_{\ell + 1}, \alpha \overset{R}{\leftarrow} Z_{p}^*, s_0=1 && \\
		\sigma_1 = (\{g^{\prod_{i\in W}s_i}\}_{W\in \mathcal{W}_{\ell, d}, g^{s_{\ell+1}}}) &&\hspace*{10pt} e(h, g^\alpha)=e(\tilde{h}, g)\\
		\sigma_2 = (\{g^{\alpha\prod_{i\in Ws_i}}\}_{W\in \mathcal{W}_{\ell, d}}, g^{\alpha s_{\ell+1}}) && : \prod_{W\in \mathcal{W}_{\ell, d}}g^{a_{W}\prod_{i\in W}s_i}g^{b{s_{\ell+1}}}\neq h\\
		\sigma = (\mathsf{bp}, \sigma_1, \sigma_2, g^{\alpha}) && \\
		\mathbb{G} \times \mathbb{G} \ni (h, \tilde{h}) \leftarrow \mathcal{A} \left(1^{\lambda}, \sigma, z\right)&& \\
		\left( a _ { 0 } , \dots , a _ { \left| \mathcal { W } _ { \ell , d } \right| } , b \right) \leftarrow \mathcal { E } \left( 1 ^ { \lambda } , \sigma , z \right) &&
		\end{aligned}\right] \le \negl{(\lambda)}\]
	}
\end{assumption}



\subsection{Interactive Proofs and Zero-knowledge Arguments}

\paragraph{Interactive proofs.} An interactive proof allows a prover $\P$ to convince a verifier $\V$ the validity of some statement. The interactive proof runs in several rounds, allowing $\V$ to ask questions in each round based on $\P$'s answers of previous rounds. We phrase this in terms of $\P$ trying to convince $\V$ that $f(x)=1$. The proof system is interesting only when the running time of $\V$ is less than the time of directly computing the function $f$. We formalize interactive proofs in the following:	
\begin{definition}\label{def:ip}
	Let $f$ be a Boolean function. A pair of interactive machines $\langle\mathcal{P}, \mathcal{V}\rangle$ is an interactive proof for $f$ with soundness $\epsilon$ if the following holds:
	\begin{itemize}
		\item {\bf Completeness.} For every $x$ such that $f(x) = 1$ it holds that $\Pr[\langle\mathcal{P}, \mathcal{V}\rangle(x)=accept]=1$.
		\item {\bf $\epsilon$-Soundness.} For any $x$ with $f(x) \neq 1$ and any $\mathcal{P}^*$ it holds that $\Pr[\langle\mathcal{P^*},\mathcal{V}\rangle=accept] \le \epsilon$
	\end{itemize}
\end{definition}


\paragraph{Zero-knowledge arguments.} An argument system for an NP relationship $R$ is a protocol between a computationally-bounded prover $\P$ and a verifier $\V$. At the end of the protocol, $\V$ is convinced by $\P$ that there exists a witness $w$ such that $(x; w) \in R$ for some input $x$. We focus on arguments of knowledge which have the stronger property that if the prover convinces the verifier of the statement validity, then the prover must know $w$. We use $\mathcal{G}$ to represent the generation phase of the public key $\pk$ and the verification key $\vk$. Formally, consider the definition below, where we assume $R$ is known to $\P$ and $\V$.

\begin{definition}\label{def::zkp}
	
	Let $R$ be an NP relation. A tuple of algorithm $(\mathcal{G}, \mathcal{P}, \mathcal{V})$ is a zero-knowledge argument for $R$ if the following holds.
	
	\begin{itemize}
		
		\item \textbf{Correctness}. For every $(\pk, \vk)$ output by $\mathcal{G}(1^\lambda)$ and $(x, w) \in R$, 
		$$\langle \P(\pk, w), \V(\vk) \rangle(x) = \accept$$
		%\babis{should not $R$ be given as input to prover and verifier?}
		\item \textbf{Soundness}. For any PPT prover $\P$, there exists a PPT extractor $\varepsilon$ such that for every $(\pk, \vk)$ output by $\mathcal{G}(1^\lambda)$ and any $x$, it holds that
		
		$$\Pr[\langle\P(\pk), \V(\vk) \rangle(x) = \accept \wedge (x, w) \notin R | w \leftarrow \varepsilon(\pk, x)] \leq \neg(\lambda)$$
		%\babis{should not $\mathcal{G}$ appear somewhere in the soundnesss definition here?}
		\item \textbf{Zero knowledge}. There exists a PPT simulator $\S$ such that for any PPT adversary $\A$, auxiliary input $z \in \{0, 1\}^{\mathsf{poly}(\lambda)}$, $(x;w)\in R$, it holds that\\
		%{\footnotesize
		$\Pr\left[\langle\P(\pk,w),\A\rangle=\accept: (\pk,\vk)\leftarrow\mathcal{G}(1^\lambda); (x,w)\leftarrow\A(z,\pk,\vk) \right] = $\\
		$\Pr\left[\langle\S(\mathsf{trap}, z, \pk),\A\rangle=\accept:(\pk,\vk,\mathsf{trap})\leftarrow\S(1^\lambda); (x,w)\leftarrow\A(z,\pk,\vk)\right]$\\
		%}
%\babis{small fontsize looks ugly...}
		%where $=$ means perfect zero knowledge \babis{this is weird. = means = probabilities. What you probably want to say is that when the probabilities are equal then the protocol is perfect zero-knowledge}. 
		
	\end{itemize}
	We say that $(\mathcal{G},\P,\V)$ is a \textbf{succinct} argument system if the
	running time of $\V$ and the total communication between $\P$ and $\V$ (proof size) are $\mathsf{poly}(\lambda,|x|,\log|w|)$.
\end{definition}

\input{CMT}
\input{zkvpd}
